---
title: "Brown et al., CNDD and the environmental context"
author: "Alissa Brown"
date: "April 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include = FALSE}
x <- c("brms", "dplyr", "vegan", "bayesplot", "knitr")
lapply(x, require, character.only = TRUE)
load("plots_mod3.RData")  # load models and data
```

# Vegetation communities

A list of the vegetation community types retained for analyses. Classification is based on the National Vegetation Classification system <http://usnvc.org/>.

```{r}
tab <- data.frame(table(dat$commPrimaryScientific))
names(tab) <- c("Veg_type", "Frequency")
kable(tab)
```

# Species names and codes

A table with all species names and their assigned species codes.

```{r}
codes <- dat %>% select(species_name, species) %>% distinct() %>% arrange(species_name)
kable(codes)
```

# Principal Components Analysis (PCA) for soil measurements

PCA results.

```{r, message = FALSE}
soil_measures <- c('Plot', 'organic', 'sand', 'silt', 'clay', 'ph', 'exchCap', 'baseSat', 
                   'Ca_ppm', 'Mg_ppm', 'K_ppm', 'Na_ppm', 'B_ppm', 'Fe_ppm', 
                   'Mn_ppm', 'Cu_ppm', 'Zn_ppm', 'Al_ppm', 'Ca_Mg_ppm', 'density')
soil_use <- dat %>% select(soil_measures)
soil_use <- soil_use %>% distinct
soil_use <- soil_use[complete.cases(soil_use), ]
soil_use[,-1] <- sapply(soil_use[,-1], as.numeric)

# par(mfrow=c(5, 4))
# par(mar=rep(0.5, 4))
# sapply(soil_use[, -1], hist)  # look for outliers

# remove outliers
out_organic <- boxplot(soil_use$organic, plot=FALSE)$out
soil_use[,"organic"] <- ifelse(
  soil_use$organic %in% out_organic, NA, soil_use$organic)

out_exchCap <- boxplot(soil_use$exchCap, plot=FALSE)$out
soil_use[,"exchCap"] <- ifelse(
  soil_use$exchCap %in% out_exchCap, NA, soil_use$exchCap)

out_baseSat <- boxplot(soil_use$baseSat, plot=FALSE)$out
soil_use[,"baseSat"] <- ifelse(
  soil_use$baseSat %in% out_baseSat, NA, soil_use$baseSat)

out_Ca_ppm <- boxplot(soil_use$Ca_ppm, plot=FALSE)$out
soil_use[,"Ca_ppm"] <- ifelse(
  soil_use$Ca_ppm %in% out_Ca_ppm, NA, soil_use$Ca_ppm)

out_Mg_ppm <- boxplot(soil_use$Mg_ppm, plot=FALSE)$out
soil_use[,"Mg_ppm"] <- ifelse(
  soil_use$Mg_ppm %in% out_Mg_ppm, NA, soil_use$Mg_ppm)

out_Na_ppm <- boxplot(soil_use$Na_ppm, plot=FALSE)$out
soil_use[,"Na_ppm"] <- ifelse(
  soil_use$Na_ppm %in% out_Na_ppm, NA, soil_use$Na_ppm)

out_B_ppm <- boxplot(soil_use$B_ppm, plot=FALSE)$out
soil_use[,"B_ppm"] <- ifelse(
  soil_use$B_ppm %in% out_B_ppm, NA, soil_use$B_ppm)

out_Mn_ppm <- boxplot(soil_use$Mn_ppm, plot=FALSE)$out
soil_use[,"Mn_ppm"] <- ifelse(
  soil_use$Mn_ppm %in% out_Mn_ppm, NA, soil_use$Mn_ppm)

out_Cu_ppm <- boxplot(soil_use$Cu_ppm, plot=FALSE)$out
soil_use[,"Cu_ppm"] <- ifelse(
  soil_use$Cu_ppm %in% out_Cu_ppm, NA, soil_use$Cu_ppm)

out_Zn_ppm <- boxplot(soil_use$Zn_ppm, plot=FALSE)$out
soil_use[,"Zn_ppm"] <- ifelse(
  soil_use$Zn_ppm %in% out_Zn_ppm, NA, soil_use$Zn_ppm)

out_Ca_Mg_ppm <- boxplot(soil_use$Ca_Mg_ppm, plot=FALSE)$out
soil_use[,"Ca_Mg_ppm"] <- ifelse(
  soil_use$Ca_Mg_ppm %in% out_Ca_Mg_ppm, NA, soil_use$Ca_Mg_ppm)

soil_use <- soil_use[complete.cases(soil_use), ]

# check data shape - needs to be normal
# dev.off()
# par(mfrow=c(5, 4))
# par(mar=rep(0.5, 4))
# sapply(soil_use[,-1], hist)

# log transform skewed measures
to_log <- c('organic', 'clay', 'exchCap', 'baseSat', 'Ca_ppm', 
            'Mg_ppm', 'K_ppm', 'Na_ppm', 'Mn_ppm', 'Cu_ppm', 
            'Zn_ppm', 'density')
foo <- which(names(soil_use) %in% to_log)

soil_use[,-1] <- soil_use[,-1] + 1
soil_use[,foo] <- sapply(soil_use[,foo], log)

# dev.off()
# par(mfrow=c(5, 4))
# par(mar=rep(0.5, 4))
# sapply(soil_use[,-1], hist)

# scale all soil measures
soil_use[,-1] <- as.data.frame(sapply(soil_use[,-1], scale))

# dev.off()
# par(mfrow=c(5, 4))
# par(mar=rep(0.5, 4))
# sapply(soil_use[,-1], hist) # looks good

# perform PCA
soil_use <- soil_use %>% ungroup()
soil_pca <- soil_use %>% select(-Plot)
pca <- rda(soil_pca)
par(mfrow=c(1,1), mar=rep(4,4))
biplot(pca, display = c("sites", "species"), type = c("text", "points"))

pca_summ <- summary(pca)
pca_summ$cont$importance
pca_summ$species
```

# Bayesian priors

```{r}
mod3$prior[, 1:4]
```

# Bayesian parameter estimates

## Fixed effects
```{r}
fixef(mod3)
```

## Random effects
### Species intercepts
```{r}
ran <- ranef(mod3)
ran$species[, , 1]
```

### Species responses to total basal area
```{r}
ran <- ranef(mod3)
ran$species[, , 2]
```

### Species responses to proportion conspecific
```{r}
ran <- ranef(mod3)
ran$species[, , 3]
```

### Species responses to the interaction between elevation and total basal area
```{r}
ran <- ranef(mod3)
ran$species[, , 4]
```

### Species responses to the interaction between elevation and proportion conspecific
```{r}
ran <- ranef(mod3)
ran$species[, , 5]
```

### Species responses to the interaction between soil fertility and total basal area
```{r}
ran <- ranef(mod3)
ran$species[, , 6]
```

### Species responses to the interaction between soil fertility and proportion conspecific
```{r}
ran <- ranef(mod3)
ran$species[, , 7]
```

### Species responses to the interaction between soil texture and total basal area
```{r}
ran <- ranef(mod3)
ran$species[, , 8]
```

### Species responses to the interaction between soil texture and proportion conspecific
```{r}
ran <- ranef(mod3)
ran$species[, , 9]
```

### Species responses to the interaction between slope steepness and total basal area
```{r}
ran <- ranef(mod3)
ran$species[, , 10]
```

### Species responses to the interaction between slope steepness and proportion conspecific
```{r}
ran <- ranef(mod3)
ran$species[, , 11]
```

### Species responses to the interaction between East-West direction and total basal area
```{r}
ran <- ranef(mod3)
ran$species[, , 12]
```

### Species responses to the interaction between East-West direction and proportion conspecific
```{r}
ran <- ranef(mod3)
ran$species[, , 13]
```

### Species responses to the interaction between North-South direction and total basal area
```{r}
ran <- ranef(mod3)
ran$species[, , 14]
```

### Species responses to the interaction between North-South direction and proportion conspecific
```{r}
ran <- ranef(mod3)
ran$species[, , 15]
```

# Posterior distributions and trace plots for model parameters

```{r}
plot(mod3, N = 3)
```

